### 2023-09-26 SujetStage/2022-pNetsubjectM2.tex
open automaton + session types

The composition of open automata refers to filling holes, it requires to check the compatibility of their sorts. Nowadays, we simply verify the compatibility of two automata to be composed based on   action labels and  arity. But, 
 we are interested in enriching holes with description of behaviours to give them enough power and enable us to check equivalences for these open systems. The description will characterise all the
interactions and the series of inputs and outputs of data exchanged between the automata and its surroundings. Precisely, the theory of session types and behavioural types  \cite{Fantechi:MLTCDP2019} are intended  to describe dynamic aspects of the behaviour of processes.

@@

The main objective of the internship is to deal with the characterization of interactions  of open automaton with its environment. These interactions can be characterised as behavioural types.
More specifically, the internship would follow the following steps:
\begin{itemize}
\item Familiarization with technical material: the semantics of  open automata and the notion behavioural types \cite{Hans:2016} that allows to describe the dynamic aspects of processes. In particular the notion of session types that focus on process interactions should be studied.

@@

\item Application to the proof of absence of deadlock upon composition.

\item  We additionally have defined a notion of bisimulation between open automata. One additional objective of the internship could be to prove that behavioural types behave well relatively to bisimilarity.

@@

The internship can be tackled only theoretically on paper, or formalised in an interactive theorem prover, or integrated with existing tools for open automata bisimilarity (based on SMT solvers).

>>>>

The composition of open automata refers to filling holes, it requires to check the compatibility of their sorts. Nowadays, we simply verify the compatibility of two automata to be composed based on   action labels and  arity (comparing actions exposed by one LTS with actions expected to fill the hole). But, 
 we are interested in enriching holes with description of behaviours; this description should look like a communication protocol. Precisely, the theory of session types and behavioural types  \cite{Fantechi:MLTCDP2019} are intended  to describe dynamic aspects of the behaviour of processes.

@@

The main objective of the internship is to deal with the characterization of interactions  of an automaton with its environment. These interactions can be characterised as behavioural types.
More specifically, the internship would follow the following steps:
\begin{itemize}
\item Familiarization with technical material: the semantics of  open automata and the notion behavioural types \cite{Hans:2016} that allows the description of the dynamic aspects of processes. In particular the notion of session types that focus on process interactions should be studied.

@@

\item Prove absence of deadlock upon composition.

\item  We  have defined a notion of bisimulation between open automata. 
One additional objective of the internship could be to prove that behavioural types behave well relatively to bisimilarity.

@@

The internship can be tackled only theoretically on paper, or formalised in an interactive theorem prover, or integrated with existing tools for open automata based on SMT solving.
